\huge A Usability Evaluation of Tor Launcher

### 2017-03-20 pets2017-paper/pets2017-paper.tex
Edit up to section 5
P.S. This is the 1000th commit!

\section{Introduction}
Tor~\cite{dingledine2004tor} is an anonymity network that routes Internet traffic through a series of relays 
to make it difficult to observe sources and destinations. Tor Browser~\cite{torbrowser} is a modified Firefox browser that routes traffic through Tor, thereby simplifying the process of configuring Tor to use existing browsers. Although Tor Browser was originally designed for anonymity, many now use Tor Browser to circumvent Internet censorship. Many countries block Tor relays specifically to prevent their citizens from circumventing censorship~\cite{winter2012great}. Tor Browser is bundled with a graphical configuration interface, Tor Launcher, which allows users to bypass censorship by configuring various bridges, proxies, and transport protocols prior to connecting to Tor. Of course, these options are only effective if users understand how to use them. Thus, we performed a usability study to examine the effectiveness of Tor Launcher and how it could be improved.

Improving Tor launcher has the obvious benefit of helping censored users connect to Tor, but this also benefits other Tor users since an increase in the overall number of users increases the anonymity set~\cite{dingledine2006anonymity}). Tor does not collect user interaction data from its real users, so we believe that usability testing research is critical to help users in heavily censored regimes and increase user adoption. In this paper, we: \\

\begin{itemize}
\item{Perform a qualitative and quantitative usability evaluation of Tor Launcher 5.0.3  under simulated censorship environments to determine why and how much it is or is not usable.}
\item{Contribute design changes that statistically significantly improve users' likelihood of connecting to Tor and also statistically significantly reduces the amount of time users take to connect.} 
\item{Provide recommendations beyond design changes to improve Tor Launcher, such as leveraging user input (requires research) and building an automated connection scheme (requires Tor-infrastructure work).}
\end{itemize}

Specifically, we describe how we started our usability evaluation by inspecting the interface (Section~\ref{sec:inspection}) and observing participants interact with it (Section~\ref{sec:qualitative}). Based on the insights from our observations, we proposed several changes to the interface (Section~\ref{sec:design}) and measured how successfully and quickly users connect to Tor before and after changes (Section~\ref{sec:quantitative}). Finally, we provide recommendations based on what we learned (Section~\ref{sec:recommendations}), we address limitations of our evaluation (Section~\ref{sec:limitations}), and frame our study in the context of existing work (Section~\ref{sec:related}). We hope that our usability evaluation serves as an informative case study and provides guidance on further improving Tor Launcher, as well as other privacy-enhancing technologies.

\section{Technical Background}
\label{sec:background}
This section discusses the network components are involved in connecting to Tor and how users can connect to Tor in various ways. An understanding of the network components or the various connection options are not required to understand our work, but we refer to these terms and methods throughout the paper.

@@

The word ``proxy'' has a specific meaning in the context of a Tor connection.
Generally, a proxy is any computer or service that allows users to make indirect connections to other services. In this sense, any Tor relay and the Tor circuit can also be considered to be a proxy. However, in this paper, the word ``proxy'' refers to a SOCKS or HTTP proxy that bypasses local firewalls.

@@

Bridges are unlisted alternative entry nodes used to connect to Tor when Tor relays are censored.
Bridges also receive packets from the user and forward traffic to middle relays, but may also run a pluggable transport (referred to as ``transport'' for shorthand). Pluggable transports specify alternative protocols by which Tor Browser may connect to the Tor network, thereby disguising Tor traffic to make it harder to detect and block bridges.
When a bridge is used, it takes the place of the guard relay.
For convenience, bridges are grouped by which transport they run.

@@

Most bridges can be blocked by blocking their static IP addresses. These include bridges that run the ``fte'' and ``fte-ipv6''~\cite{fte} pluggable transports, which disguise the Tor protocol as another protocol (such as HTTP); these also include the ``obfs3''~\cite{obfs3}, ``obfs4''~\cite{obfs4}, and ``scramblesuit''~\cite{scramblesuit} transports, which encrypt or alter the Tor protocol to appear as random noise. Other bridges are less susceptible to blocking because of how they work. Bridges that run the ``flashproxy''~\cite{flashproxy} and ``meek''~\cite{fifield2015blocking} transports route traffic through third party web browsers and content delivery networks. The bridges and transports that are accessible in each country vary. Figure~\ref{fig:old-bridge-combobox} shows the bridge and transport options at the time of our study.

@@

\subsection{Connecting to Tor} 
It is important to note that there are many valid configuration settings that can be used to connect to the Tor network. A user may or may not need a combination of proxy and bridge.
A user who does not need a bridge or proxy can still connect with a bridge, proxy, or both, provided that they are configured correctly (though adding unnecessary network hops or obfuscation protocols may slow the connection down).

Users can configure a bridge by choosing a transport, which selects the group of bridges that use that transport. For instance, choosing the ``obfs3'' option in Figure~\ref{fig:old-bridge-combobox} will select a group of bridges that perform the obfs3 transport protocol as okay to use, and the user will be assigned one of these bridges as their guard relay. If these built-in options do not work, users can obtain bridge configurations through out-of-band channels~\cite{bridgedb}. Configuring a proxy requires entering in the proxy protocol, IP address, port, and additional optional fields (Figure~\ref{fig:old-proxy}). The user must locate this information.

@@

\section{Methodology} 
This section discusses the usability testing process used to evaluate Tor Launcher, the simulated censorship environments we used to require users to configure different network components to connect to Tor, and what our evaluation metrics  we collected to deem the interface usable or not usable.

\subsection{Testing Pipeline} 
We began by performing a cognitive walkthrough~\cite{wharton1994cognitive,cognitive-walkthrough} to inspect the interface to anticipate potential problems and formally assess its pros and cons. A cognitive walkthrough is a formal method of inspection performed by researchers without any user involvement. We then conducted qualitative and quantitative usability tests. We believe both methods were necessary for a complete usability evaluation. Qualitative testing tells you {\it why or how} it is or is not usable, while quantitative testing tells you {\it how much} it is or is not usable. We used our observations from the qualitative study to design interface changes that would address common issues that these participants encountered. We used our quantitative testing setup to measure the usability of our proposed changes, as compared to the original interface.

We performed our qualitative experiments one participant at a time, so that we could observe their behaviors in real time and interview them about their experience after use. We recruited five participants per experimental condition to gather qualitative data; this is considered the optimal number of measurements to maximize the cost-benefit ratio~\cite{howmanyusers}. We performed our quantitative experiments at a behavioral laboratory many participants at a time, using instrumentation to capture evaluation metrics. We recruited twenty users per experimental condition to gather quantitative data; this is considered to be the minimum for yielding statistically significant measurements~\cite{howmanyusers}.

@@

In the first environment, E1, particular websites were blocked, which meant that they could be accessed via a default Tor configuration (i.e., bridges or proxies were not needed). In the second environment, E2, we additionally blocked the IP addresses of public Tor relays, which meant that Tor could still be accessed by enabling any of the included bridges. Finally, the third environment, E3, additionally blocked several of the included bridges, which required experimenting with several different bridges---beyond the recommended option---in order to successfully connect.

@@

\begin{itemize}
\item {\bfseries Completion rate:}  the percentage of users that want to connect to Tor and successfully do so. The higher the completion rate, the better. The ideal case would be that everyone is able to connect to Tor if they want to. Even if users do not successfully connect to Tor at first but eventually do so during the experiment, they are considered to have successfully connected to Tor for this metric.
\item {\bfseries Connection time:} the amount of time users take to connect to Tor. The less time it takes for users to connect, the better. The ideal case would be that the connection time is very close to the time it takes for the application to communicate with the servers and establish a circuit of relays for the connection, which can take up to a couple minutes. Connection time includes this bootstrapping time and any additional time that people spend in the interface. 
\item {\bfseries Configuration time:} the amount of time users spend in the interface. The less time it takes for users to configure their preferences, the better. Configuration time is included in connection time, so it will always be less than connection time. We defined configuration time as any time spent on the bridge and proxy screens, since users were most likely configuring bridges and proxies. We excluded any time spent on the progress screen, because users were most likely waiting for connections to establish or timeout and not necessarily interacting with the interface. This should be as short as possible.
\end{itemize}

Completion rate is the fundamental usability metric that measures if users can accomplish the task. We defined success as a binary metric that is true if a user connected to Tor and false if a user did not. We consider this to be the most important usability metric, as this directly measures if users could connect to Tor.

Task time is a supplemental usability metric that measures user productivity during the task. We measured this in two ways. Connection time measures how long users took to connect to Tor, while configuration time to measures how long users spent actively configuring bridges and proxies. We use both metrics to quantify productivity since the bootstrapping time, and hence, time spent waiting for the connection to succeed or fail, can dominate the measurement. Connection time is always greater than the configuration time, since it includes the configuration time. We consider connection time and configuration time to be equally important in measuring user productivity.

@@

\begin{itemize}
\item {\bfseries Technical terms:} The interface uses technical and Tor-specific terms that are likely unfamiliar to an average user (e.g., ``Tor network,'' ``local proxy,'' ``bridges,'' and ``transports''). 
\item {\bfseries Too many choices:}  Users who do not need a bridge or proxy are able to configure them. There are nine choices for bridges by transport and more options to input a custom bridge. Almost any HTTP or SOCKS proxy can be used as the proxy. Custom bridge lines require the bridge information with specific syntax, but this syntax is not specified in the interface. Proxy fields also do not provide syntactic guidelines. See the whole list of possible errors we identified in Appendix~\ref{states}. 
\item {\bfseries Poor feedback:} Even if there was a syntactic error in the bridge or proxy fields, the user is not notified until after a connection is attempted and fails. If a censored bridge was chosen, the connection timeout takes several minutes. Error messages were often technical and did not give suggestions.
\end{itemize}

@@

\section{Qualitative Testing}
\label{sec:qualitative}
We performed our qualitative experiments one participant at a time, so that we could observe their behaviors in real time and interview them about their experience after use. We observed user behaviors to identify Tor stumbling points and then interviewed them afterwards to gather explanatory data and their opinions.

>>>>

\section{Introduction}
Tor~\cite{dingledine2004tor} is an anonymity network that routes Internet traffic through a series of relays 
to make it difficult to observe sources and destinations.
Tor Browser~\cite{torbrowser} is a modified Firefox browser that routes traffic through Tor, designed to minimize the configuration required before connecting.
Although Tor was originally designed for anonymity, many now use it to circumvent Internet censorship.
Because of this, many countries block Tor relays specifically to prevent their citizens from circumventing censorship~\cite{winter2012great}.

Tor Browser is bundled with a graphical configuration interface, Tor Launcher, which allows users to bypass censorship by configuring various bridges, proxies, and transport protocols prior to connecting to Tor.
Of course, these options are only effective if users understand how to use them.
To understand to what extent this is the case, we researched the effectiveness of Tor Launcher as well as how it could be improved.

Improving Tor Launcher has the obvious benefit of helping censored users connect to Tor, but this also benefits other Tor users since an increase in the overall number of users increases the anonymity set~\cite{dingledine2006anonymity}). Tor does not collect user interaction data from its real users, so we believe that usability testing research is critical to help users in heavily censored regimes and increase user adoption. In this paper, we: \\

\begin{itemize}
\item{Perform a qualitative and quantitative usability evaluation of Tor Launcher 5.0.3  under simulated censorship environments}
\item{Contribute design changes that improve users' likelihood of connecting to Tor and also reduce the amount of time users take to connect}
\item{Provide recommendations beyond design changes to improve Tor Launcher, such as leveraging user input (requires research) and building an automated connection scheme (requires Tor infrastructure work).}
\end{itemize}

Specifically, we describe how the initial evaluation of the interface (Section~\ref{sec:inspection}) and our observations of participants' interactions with it (Section~\ref{sec:qualitative}).
Based on the insights from our observations, we proposed several changes to the interface (Section~\ref{sec:design}) and measured how successfully and quickly users connect to Tor before and after changes (Section~\ref{sec:quantitative}).
Finally, we provide recommendations based on what we learned (Section~\ref{sec:recommendations}), we address limitations of our evaluation (Section~\ref{sec:limitations}), and frame our study in the context of existing work (Section~\ref{sec:related}).
We hope that our usability evaluation serves as an informative case study and provides guidance on further improving Tor Launcher, as well as other privacy-enhancing technologies.

\section{Technical Background}
\label{sec:background}
This section discusses the network components involved in connecting to Tor.
An understanding of the various connection options are not required to understand our work, but we refer to these terms and methods throughout the paper.

@@

The word ``proxy'' has a specific meaning in the context of a Tor connection.
Generally, a proxy is any computer or service that allows users to make indirect connections to other services. In this sense, any Tor relay and the Tor circuit can also be considered a proxy. However, in this paper, the word ``proxy'' refers to a SOCKS or HTTP proxy that bypasses local firewalls.

@@

Bridges are unlisted alternative entry nodes used to connect to Tor when Tor relays are censored.
Bridges also receive packets from the user and forward traffic to middle relays, but may also run a pluggable transport (referred to as ``transport'' for shorthand). Pluggable transports specify alternative protocols by which Tor Browser may connect to the Tor network, thereby disguising Tor traffic to make it harder to detect and block bridges.
When a bridge is used, it takes the place of the guard relay.
For convenience, bridges are grouped in the interface by which transport they run.

@@

Most bridges can be blocked by blocking their static IP addresses. These include bridges that run the \texttt{fte} and \texttt{fte-ipv6}~\cite{fte} pluggable transports, which disguise the Tor protocol as another protocol (such as HTTP); these also include the \texttt{obfs3}~\cite{obfs3}, \texttt{obfs4}~\cite{obfs4}, and \texttt{scramblesuit}~\cite{scramblesuit} transports, which encrypt or alter the Tor protocol to appear as random noise.

Other bridges are less susceptible to blocking because of how they work. Bridges that run the \texttt{flashproxy}~\cite{flashproxy} and \texttt{meek}~\cite{fifield2015blocking} transports route traffic through third party web browsers and content delivery networks. The bridges and transports that are accessible in each country vary. Figure~\ref{fig:old-bridge-combobox} shows the bridge and transport options at the time of our study.

@@

\subsection{Connecting to Tor} 
It is important to note that there are many valid configuration settings that can be used to connect to the Tor network. A user may or may not need a proxy and/or a bridge.
A user who does not need a bridge or proxy can still connect with a bridge, proxy, or both, provided that these are configured correctly (though adding unnecessary network hops or obfuscation protocols may slow the connection down).

Users can configure a bridge by choosing a transport, which selects the group of bridges that use that transport.
For instance, a user choosing the \texttt{obfs3} option in Figure~\ref{fig:old-bridge-combobox} will be assigned one of these bridges as their guard relay.
If these built-in options do not work, users can obtain bridge configurations through out-of-band channels~\cite{bridgedb}.
Configuring a proxy requires entering in the proxy protocol, IP address, port, and additional optional fields (Figure~\ref{fig:old-proxy}).
The user must obtain this information themselves.

@@

\section{Methodology} 
This section discusses our usability testing process, including the simulated censorship environments we used to require users to configure different network components to connect to Tor and the evaluation metrics  we collected.

\subsection{Testing Pipeline} 
We began by performing a cognitive walkthrough~\cite{wharton1994cognitive,cognitive-walkthrough} to inspect the interface to anticipate potential problems and formally assess its pros and cons.
A cognitive walkthrough is a formal method of inspection performed by researchers without any user involvement.
We then conducted qualitative and quantitative usability tests.
We believe both methods were necessary for a complete usability evaluation.
Qualitative testing illuminates {\it why or how} the software is or is not usable, while quantitative testing measures {\it to what extent}.
We used our observations from the qualitative study to design interface changes that would address common issues that participants encountered.
We used our quantitative testing setup to measure the usability of our proposed changes, as compared to the original interface.

We performed our qualitative experiments one participant at a time, so that we could observe their behaviors in real time and interview them about their experience after user.
We recruited five participants per experimental condition to gather qualitative data; this is considered the optimal number of measurements to maximize the cost-benefit ratio~\cite{howmanyusers}.
We performed our quantitative experiments at a behavioral laboratory many participants at a time, using instrumentation to capture evaluation metrics.
We recruited twenty users per experimental condition to gather quantitative data; this is considered to be the minimum for yielding statistically significant measurements~\cite{howmanyusers}.

@@

In the first environment, E1, particular websites were blocked, which meant that they could be accessed using a default Tor configuration (i.e., bridges or proxies were not needed). In the second environment, E2, we additionally blocked the IP addresses of public Tor relays, which meant that Tor could still be accessed by enabling any of the included bridges. Finally, the third environment, E3, additionally blocked several of the included bridges, which required experimenting with several different bridges---beyond the recommended option---in order to successfully connect.

@@

\begin{itemize}
\item {\bfseries Completion rate:}  the percentage of users that want to connect to Tor and successfully do so. The higher the completion rate, the better. The ideal case would be that everyone is able to connect to Tor if they want to. Even if users do not successfully connect to Tor at first but eventually do so during the experiment, they are considered to have successfully connected to Tor for this metric.
\item {\bfseries Connection time:} the amount of time users take to connect to Tor. The less time it takes for users to connect, the better. The ideal case would be that the connection time is very close to the time it takes for the application to communicate with the servers and establish a circuit of relays for the connection, which can take up to several minutes. Connection time includes this bootstrapping time and any additional time that people spend in the interface.
\item {\bfseries Configuration time:} the amount of time users spend in the interface. The less time it takes for users to configure their preferences, the better. Configuration time is included in connection time, so it will always be less than connection time. We defined configuration time as any time spent on the bridge and proxy screens, since users were most likely configuring bridges and proxies. We excluded any time spent on the progress screen, because users were most likely waiting for connections to establish or timeout and not necessarily interacting with the interface. This should be as short as possible.
\end{itemize}

Completion rate is the most important usability metric, measuring if users can accomplish the task. We defined success as a binary metric that is true if a user connected to Tor and false if a user did not.
Task time is a supplemental usability metric that measures user productivity during the task. We use both connection and configuration time to quantify productivity since the bootstrapping time, and hence, time spent waiting for the connection to succeed or fail, can dominate the measurement.

@@

\begin{itemize}
\item {\bfseries Technical terms:} The interface uses technical and Tor-specific terms that are likely unfamiliar to an average user (e.g., ``Tor network,'' ``local proxy,'' ``bridges,'' and ``transports''). 
\item {\bfseries Too many choices:}  Users who do not need a bridge or proxy are able to configure them.
There are nine choices for bridges and more options to input a custom bridge.
Almost any HTTP or SOCKS proxy can be used as the proxy.
Custom bridge lines require the bridge information in a specific syntax, but this syntax is not specified in the interface.
Proxy fields also do not provide syntactic guidelines.
The full list of possible errors we identified can be found in Appendix~\ref{states}.
\item {\bfseries Poor feedback:} Even if there was a syntactic error in the bridge or proxy fields, the user is not notified until after a connection is attempted and fails. If a censored bridge was chosen, the connection timeout takes several minutes. Error messages were often technical and did not give suggestions for next steps.
\end{itemize}

@@

\section{Qualitative Testing}
\label{sec:qualitative}
We performed our qualitative experiments one participant at a time, so that we could observe their behaviors in real time identify stumbling points in the interface.
At the completion of the task, we interviewed our participants to gather explanatory data and their opinions.
